Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-30 numerical.tex
revising

Unless otherwise stated, our fiducial disk parameters are 
$(\gamma, \Gamma) = (1.4, 1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.05)$, and 
vertical domain size $\zmax=5H$. This setup is effectively vertically
isothermal since $c_s(\zmax) \simeq 0.9c_s(0)$. In Fig. \ref{omega_z}
we plot the corresponding equilibrium rotation profile, which shows
that $\Omega$ decreases slightly away from the midplane.

@@

\subsection{VSI with rapid thermal relaxation}\label{vertiso_pertiso} 
We first calculate the VSI in a disk with rapid thermal relaxation by
setting $\beta=10^{-3}$. (Smaller values yield similar results.) 
This setup mimics strictly isothermal perturbations in a vertically
isothermal disk, as considered by \cite{nelson13}, \cite{mcnally14} and
\cite{barker15}. %Note, however, that \citeauthor{nelson13} and
% \citeauthor{barker15} employed the  approximations described in \S\ref{analytical} throughout their analysis, which are relaxed in our
% numerical study.  
For completeness, and as a way of introducing terminology, 
we review this case before considering the effect of 
thermal relaxation. A more detailed analysis is given by \cite{barker15}.

@@

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z5}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z7} 
  \caption{Maximum VSI growth rates in the fiducial disk
     model as a function of the thermal relaxation timescale
     $\beta$. The vertical line is the
     critical thermal timescale $\beta_\mathrm{crit}$ obtained  
     from Eq. \ref{iso_vsi_cond}. 
     \label{bcrit_compare1}}   
 \end{figure}

@@

Here we briefly consider a vertically non-isothermal disk model with  
$(p,q,\epsilon)=(0,-1,0.05)$ and $\Gamma=1.0837$. This value of 
$\Gamma$ is chosen so that the standard vertical domain size $\zmax =
5H$ effectively corresponds to the disk surface ($\zmax\simeq
H_s$). The disk temperature decreases rapidly with height such that 
$c_s(\zmax)\simeq 10^{-4}c_s(0)$.

>>>>

Unless otherwise stated, our fiducial disk parameters are 
$(\gamma, \Gamma) = (1.4, 1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.05)$, and 
vertical domain size $\zmax=5H$. This setup is effectively vertically
isothermal since $c_s^2(\zmax) \simeq 0.9c_s^2(0)$. In Fig. \ref{omega_z}
we plot the corresponding equilibrium rotation profile, which shows
that $\Omega$ decreases slightly away from the midplane.

@@

\subsection{VSI with rapid thermal relaxation}\label{vertiso_pertiso} 
We first calculate the VSI in a disk with rapid thermal relaxation by
setting $\beta=10^{-3}$. (Smaller values yield similar results.) 
This setup mimics strictly isothermal perturbations in a vertically
isothermal disk, as considered by \cite{nelson13}, \cite{mcnally14} and
\cite{barker15}. %Note, however, that \citeauthor{nelson13} and
% \citeauthor{barker15} employed the  approximations described in \S\ref{analytical} throughout their analysis, which are relaxed in our
% numerical study.  
For completeness, and as a way of introducing terminology, 
we review this case before considering the effect of 
thermal relaxation. %A more detailed analysis is given by \cite{barker15}.

@@

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z3}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z5}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z7} 
  \caption{Maximum VSI growth rates in the fiducial disk 
     model as a function of the thermal relaxation timescale
     $\beta$, for $\zmax=3H$ (top), $\zmax=5H$ (middle, standard
     setup) and $\zmax=7H$ (bottom). The vertical line is the
     critical thermal timescale $\beta_\mathrm{crit}$ obtained  
     from Eq. \ref{iso_vsi_cond}. 
     \label{bcrit_compare1}}   
 \end{figure}

@@

Here we briefly consider a vertically non-isothermal disk model with  
$(p,q,\epsilon)=(0,-1,0.05)$ and $\Gamma=1.0837$. This value of 
$\Gamma$ is chosen so that the standard vertical domain size $\zmax =
5H$ effectively corresponds to the disk surface ($\zmax\simeq
H_s$). The disk temperature decreases rapidly with height such that 
$c_s^2(\zmax)\simeq 10^{-4}c_s^2(0)$.
